I wish to express my gratitude to my advisor, Prof. K. Madhava Krishna from the Robotics Research Centre at IIIT Hyderabad, for being a huge pillar of support, from helping me recognize my aptitude for research to allowing me to hone my skills in multiple sub-domains under Robotics. I am also immensely thankful to Prof. Arun Kumar Singh from the University of Tartu, Estonia, for guiding me in every step of my research journey, starting from formulating research problems to hands-on pair programming. I would also like to thank Prof. Ponnurangam Kumaraguru for boosting my confidence in research paper writing and allowing me to explore diverse domains during my Independent Study under his supervision.

Next, I would like to thank the faculty at IIIT Hyderabad for providing rigorous training in mathematics and Electronics at the undergraduate level that helped me build my foundations in Robotics. I owe my understanding of Robotics to the project-based electives taught by Prof. Harikumar Kandath, Prof. K. Madhava Krishna, and Prof. Spandan Roy and to the Robotics Summer School organized by my seniors. 


I would also like to convey my gratitude to Fatemeh Rastgar, a Junior Ph.D. Researcher under Prof. Singh, who has supported me as a mentor, helping me gain a strong understanding of multi-robot planning algorithms and GPU-acceleration libraries. My teammates in the manipulator project from RRC - Vishal Reddy Mandadi, Md. Nomaan Qureshi, Bipasha Sen, Aditya Aggarwal, and Kallol Saha have also played a significant role in my understanding of physics simulators and Reinforcement Learning-based robot control. My student mentors - Mithun Babu Nallana, Jyotish P, and Raghu Ram Theerthala, helped me take my first steps in robot path planning at the end of my second year and patiently reviewed my implementations of state-of-the-art algorithms. In addition, my batchmates and fellow research students at RRC have often readily contributed suggestions to improve my research work and validated my ideas. Aditya Sharma Meduri has helped me get acquainted with ROS and Gazebo to run simulations for my paper. My batchmates and friends, Vedant Mundheda, Karan Mirakhor, Rahul Swayampakula, Nipun Wahi and Manav Bhatia have constantly offered their feedback on my research projects as teammates in course projects and external competitions. They have also supported me throughout my degree, helping me balance my academics with research, along with my social life on campus.

Finally, I would like to thank my mother and grandmother for constantly supporting me and believing in me. 
